Sequential Composition for Relaxed Memory: Pomsets with Predicate Transformers

### 2021-06-29 paper/arm.tex
small cleanup of arm semantics

\section{Lowering \PwTmcaTITLE{} to Arm}
\label{sec:arm}
For simplicity, we restrict to top level parallel composition.
% and ignore
% fences\footnote{Fences are not actions in \armeight{}, which complicates the
%   theorem statements.%
%   \app{Is it true? AFAIU, it isn't, they are explicitly used in definition of $\textsc{bob}$.
%     I've just found that Jade prefers to omit them in pictures.}}.
% JWR: I didn't know that!  It's better with them...

\subsection{Arm executions}
\begin{definition}
  An \emph{\armeight{} execution graph}, $\aEG$, is tuple
  $(\aEvs, \labeling, {\rpoloc}, {\rlob})$ such that
  \begin{enumerate}[,label=(\textsc{a}\arabic*),ref=\textsc{a}\arabic*]
  \item $\aEvs\subseteq\AllEvents$ is a set of {events},
  \item $\labeling: \aEvs \fun \Act$ defines a {label} for each event,
  \item ${\rpoloc} : \aEvs\times\aEvs$, is a per-thread, per-location total
    order, capturing \emph{per-location program order},
  \item ${\rlob} : \aEvs\times\aEvs$, is a per-thread partial order capturing
    \emph{locally-ordered-before}, such that
    \begin{enumerate}
    \item \label{arm-lob-poloc}
      ${\rpoloc} \cup {\rlob}$ is acyclic.
    \end{enumerate}
  \end{enumerate}
\end{definition}

@@

\begin{definition}
  Execution $\aEG$ is
  \emph{$({\rco}, {\rrfx}, {\rgcb})$-valid}, under \emph{External Global
    Consistency} (\EGC{}) if
  \begin{enumerate}[label=(\textsc{a}\arabic*),ref=\textsc{a}\arabic*]
    \setcounter{enumi}{4}
  %\item[\eqref{arm-co}] and \eqref{arm-rf}, as for \EC,
  \item \label{arm-co}
    ${\rco} : \aEvs\times\aEvs$, is a per-location total order on
    writes, capturing \emph{coherence}, 
  \item \label{arm-rf}
    ${\rrfx} : \aEvs\times\aEvs$, is a surjective and injective
    relation on reads, capturing \emph{reads-from}, such that
    \begin{enumerate}
    \item \label{arm-match}
      if $\bEv\xrfx\aEv$ then $\labeling(\bEv) \rmatches \labeling(\aEv)$,      
    \item \label{arm-local}
      ${\rpoloc} \cup {\rco} \cup {\rrfx} \cup {\rfr}$ is acyclic,
      where $\aEv\xfr\cEv$ if %$(\exists\bEv)$
      $\aEv\xrfxinv\bEv\xco\cEv$, for some $\bEv$,
      % \stepcounter{enumi}
      % \item[] \labeltext[\textsc{a}2]{}{arm-cb}
    \end{enumerate}
    % \item ${\rpoloc} \cup {\rco} \cup {\rrfx} \cup {\rfr}$ is acyclic, where
    %   $\bEv\xfr\aEv$ if $(\exists\cEv)$ $\bEv\xrfxinv\cEv\xco\aEv$,
  \item \label{arm-gcb}
    ${\rgcb}\supseteq\PBR{{\rco}\cup{\rrfx}}$ is a linear order %, capturing \emph{globally completes before}, %$\rcb : % \aEvs\times\aEvs$
    such that 
    \begin{enumerate}%[leftmargin=0pt]
      % \item if $\bEv\xco\aEv$ then $\bEv\xgcb\aEv$,
      % \item if $\bEv\xrfx\aEv$ then $\bEv\xgcb\aEv$,
    \item \label{arm-gcb-blocks}
      if $\bEv\xrfx\aEv$ and $\labeling(\cEv) \rblocks \labeling(\aEv)$ then either $\cEv\xgcb\bEv$ or $\aEv\xgcb\cEv$, 
    \item \label{arm-gcb-lob}
      if $\aEv\xlob\cEv$ then either $\aEv\xgcb\cEv$ or $(\exists\bEv)$
      $\bEv\xrfx\aEv$ and $\bEv\xpoloc\aEv$ but not $\bEv\xlob\cEv$.
    \end{enumerate}
  \end{enumerate}

>>>>

\section{Lowering \PwTmcaTITLE{} to Arm}
\label{sec:arm}
For simplicity, we restrict to top-level parallel composition.
% and ignore
% fences\footnote{Fences are not actions in \armeight{}, which complicates the
%   theorem statements.%
%   \app{Is it true? AFAIU, it isn't, they are explicitly used in definition of $\textsc{bob}$.
%     I've just found that Jade prefers to omit them in pictures.}}.
% JWR: I didn't know that!  It's better with them...

\subsection{Arm executions}
\begin{definition}
  An \emph{\armeight{} execution graph}, $\aEG$, is tuple
  $(\aEvs, \labeling, {\rpoloc}, {\rlob})$ such that
  \begin{enumerate}[,label=(\textsc{a}\arabic*),ref=\textsc{a}\arabic*]
  \item $\aEvs\subseteq\AllEvents$ is a set of {events},
  \item $\labeling: \aEvs \fun \Act$ defines a {label} for each event,
  \item ${\rpoloc} \subseteq \aEvs{\times}\aEvs$, is a per-thread, per-location total
    order, capturing \emph{per-location program order},
  \item ${\rlob} \subseteq \aEvs\times\aEvs$, is a per-thread partial order capturing
    \emph{locally-ordered-before}, such that
    \begin{enumerate}
    \item \label{arm-lob-poloc}
      ${\rpoloc} \cup {\rlob}$ is acyclic.
    \end{enumerate}
  \end{enumerate}
\end{definition}

@@

\begin{definition}
  Execution $\aEG$ is
  \emph{$({\rco}, {\rrfx}, {\rgcb})$-valid}, under \emph{External Global
    Consistency} (\EGC{}) if
  \begin{enumerate}[label=(\textsc{a}\arabic*),ref=\textsc{a}\arabic*]
    \setcounter{enumi}{4}
  %\item[\eqref{arm-co}] and \eqref{arm-rf}, as for \EC,
  \item \label{arm-co}
    ${\rco} \subseteq \aEvs\times\aEvs$, is a per-location total order on
    writes, capturing \emph{coherence}, 
  \item \label{arm-rf}
    ${\rrfx} \subseteq \aEvs\times\aEvs$, is a relation, capturing \emph{reads-from}, such that
    \begin{enumerate}
    \item \label{arm-rf-reads}
      ${\rrfx}$ is surjective and injective relation on $\{\aEv\in\aEvs\mid\labeling(\aEv)\;\text{is a}\;\sread\}$,
    \item \label{arm-match}
      if $\bEv\xrfx\aEv$ then $\labeling(\bEv) \rmatches \labeling(\aEv)$,      
    \item \label{arm-local}
      ${\rpoloc} \cup {\rco} \cup {\rrfx} \cup {\rfr}$ is acyclic,
      where $\aEv\xfr\cEv$ if %$(\exists\bEv)$
      $\aEv\xrfxinv\bEv\xco\cEv$, for some $\bEv$,
      % \stepcounter{enumi}
      % \item[] \labeltext[\textsc{a}2]{}{arm-cb}
    \end{enumerate}
    % \item ${\rpoloc} \cup {\rco} \cup {\rrfx} \cup {\rfr}$ is acyclic, where
    %   $\bEv\xfr\aEv$ if $(\exists\cEv)$ $\bEv\xrfxinv\cEv\xco\aEv$,
  \item \label{arm-gcb}
    ${\rgcb}\supseteq\PBR{{\rco}\cup{\rrfx}}$ is a linear order %, capturing \emph{globally completes before}, %$\rcb : % \aEvs\times\aEvs$
    such that 
    \begin{enumerate}%[leftmargin=0pt]
      % \item if $\bEv\xco\aEv$ then $\bEv\xgcb\aEv$,
      % \item if $\bEv\xrfx\aEv$ then $\bEv\xgcb\aEv$,
    \item \label{arm-gcb-blocks}
      if $\bEv\xrfx\aEv$ and $\labeling(\cEv) \rblocks \labeling(\aEv)$ then either $\cEv\xgcb\bEv$ or $\aEv\xgcb\cEv$, 
    \item \label{arm-gcb-lob}
      if $\aEv\xlob\cEv$ then either $\aEv\xgcb\cEv$ or $(\exists\bEv)$
      $\bEv\xrfx\aEv$ and $\bEv\xpoloc\aEv$ but not $\bEv\xlob\cEv$.
    \end{enumerate}
  \end{enumerate}
